### 2016-11-23 manuscript.tex
20161123

%\documentclass[draftclsnofoot,onecolumn]{IEEEtran}
\documentclass[3p]{elsarticle}
\usepackage{tipa}
\usepackage{amsmath}%math
\usepackage{amsthm}%proof
\usepackage{xcolor}%color
\usepackage{bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{textcomp}
\usepackage[colorlinks,linkcolor=black,anchorcolor=black,citecolor=black]{hyperref}
\usepackage{amsfonts,amssymb}
\usepackage{color,soul}
\theoremstyle{plain}
\newtheorem{myas}{Assumption}
\newtheorem{mydef}{Definition}
\newtheorem{mylem}{Lemma}
\newtheorem{mythm}{Theorem}
\newtheorem{mypro}{Properties}
%\theoremstyle{remark}
\theoremstyle{remark}
\newtheorem{myrem}{Remark}
\begin{document}
\begin{frontmatter}
\title{Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System}
\author{Zhiqiang Ma}
\author{Guanghui Sun\corref{cor1}}
{\author{Zhengkai Li\corref{cor2}}}
\ead{guanghuisun@hit.edu.cn}
\cortext[cor1]{Corresponding author}
\address{Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China}

\begin{abstract}
This paper presents a novel adaptive dynamic sliding mode scheme for the deployment of the tethered satellite system. To overcome the limited inputs of the dynamic model, a new scalable dimensionless transformation and strictly bounded terms are designed, which guarantee the existence of the ratio between limited inputs and command signals. The adaptive saturated sliding mode control scheme is proposed to govern the tether deployment by introducing the adaptive rate into the system gain. The stability analysis indicates that this measure can eliminate the uncertainty caused by the input limitation and guarantee the asymptotically stability of the deployment dynamics. Finally, numerical simulations are presented to verify the effectiveness of the proposed control laws.
\end{abstract}
\begin{keyword}
Tethered satellite system;  Sliding mode control;  Input limitation; Adaptive control
\end{keyword}
\end{frontmatter}
\section{Introduction}
The deployment is the key technology of the space tethered satellite system (TSS)~\cite{wen2008advances,yousefian2015anti}, which supports various applications for space explorations and scientific researches, such as space debris removals, space elevators, satellite de-orbits and the tethered space robots~\cite{zhao2014thrust,kojima2015mission,zhong2013dynamics,zhong2013long,huang2014optimal,Huang2016Impact,Huang201637,Huang2016}. Many international space research institutes have launched orbital, suborbital and ground experiments, such as TSS-1R, SEDS, MAST and YES2, in an effort to verify the feasibility of the proposed theoretical and simulation schemes for the deployment of the TSS~\cite{williams2012review,robitaille2006interpreting,stone1998tss,williams2009yes2}. These experimental results intensively help researchers proceed to promote advanced mechanics, kinetic and control studies, to tackle the future challenges of space activities~\cite{yu2016nonlinear,wen2015space,aslanov2016swing,meng2016lyapunov,zhang2015line,Xu2016Hybrid}.

Recent scientific researches about deployments of the TSS focus on the stabilization of the dynamics, and have provided different kinds of control schemes, including tension, thruster, electrodynamic and Coulomb approaches~\cite{kojima2015stabilisation,qi2016dynamics,zhong2016research}. Among these stabilization strategies, the hybrid tension and thruster control technique has been a great boost to new developments of the deployment of the TSS. Iki et al. implemented the ground deployment experiment, and utilized the thruster to conduct the deployment of a kilometers-length electrodynamic tether from a special spool-type reel~\cite{iki2014experiments}. Aslanov et al. presented a novel model about the spatial motion of a large passive satellite and the tug connected by space tether, and the motion of the overall system was excited by a thruster force acting on the tug~\cite{aslanov2013dynamics}. Huang~et~al. investigated the attitude and deployment dynamics of the tethered space robot from the standpoint of thruster consumption~\cite{huang2015coupling}. Jung et al. adopted the thruster scheme to achieve the ideal deployment rate during the deployment of three-body tethered satellites~\cite{jung2015nonlinear}. Liu et al. equipped the subsatellite with the thruster system, consisting of four pairs of thrusters, in order to regulate the attitude and deployment dynamics~\cite{liu2014attitude}. Zakrzhevskii divided the deployment process into two stages: in the first stage, the thruster was used to propel the subsatellite to the desired starting point of the next stage, and in the second stage, the tether was adjusted to the stable vertical configuration using tension generated by friction~\cite{zakrzhevskii2016method}. Mantellato et al. reported the thruster-aided deployment of tape tethers driven by both the tension and thruster forces~\cite{mantellato2015thrust}. These hybrid strategies has several advantages, such as simplifying the input model greatly, non-environmental dependence (magnetic field dependence), easy assembly and physical implementation, and thus is broadly utilized in engineering applications compared with other control strategies.

However, there also exist some obstacles in the hybrid control design. Particularly, it's a prime problem that tension actuators and thrusters are both constrained by physical and chemical conditions, i.e., the space tether is not available to be repressed like the stick, and meanwhile too big tether length rate might break the tether. Also, the thruster can not provide the infinite propelling force~\cite{martinez1998spacecraft}. In simpler terms, the tension and the thruster force for the deployment are both limited, and in the view of control theory, the phenomena usually causes the input limitation of a dynamic system. Substituting limited inputs into unmodified conventional control scheme designed for systems with unlimited inputs results in poor performance or instability, and theoretically, such kind of implement is unreasonable. Hence to cope with the issues mentioned above, many scientific researchers spent much effort on the issue about the input limitation, providing many solutions on satellite arrangements and deployments. Huang et al. proposed an adaptive postcapture backstepping control law for the tethered space robot, and analyzed the effect from constrained inputs by using specific auxiliary design systems~\cite{huang2015adaptive}. Wen et al. took into account the positive tension constraint and designed the feedback control with the special saturation technique~\cite{wen2016constrained}. Wen also proposed some remarkable achievements about deployment dynamics using the optimal control theory~\cite{wen2008optimal}. Liu et al. reported the variable structure scheme for the deployment with the special structural thruster system, where the thrusters were used to ensure the positive tension during the mission~\cite{yingying2012variable}. In summary, existing studies deal with the major issue about the deployment with restricted tension and thruster force in two ways, accordingly, the suitable physical compensation and the elaborate design of limited input. In aspect of designing elaborate limited inputs, as Wen et al. explained, the limited terms, including the tension and the states of deployment dynamics, can be viewed as seeking the solution of a nonlinear optimal control issue~\cite{wen2015tension}. Hu reported an adaptive sliding mode control law with the input limitation by using the reduced sliding dynamic design approach~\cite{Hu2009Robust}, which requires an assumption that the command signal is bounded strictly, and for actual physical systems, this assumption always exits. In this paper, a new dynamic adaptive saturated sliding mode  control for the deployment of the TSS is studied. For coping with the input limitation in the hybrid control strategy, the strictly saturated command signal is constructed with Sine saturation function. Based on the feature of the saturated command signal, an adaptive rate is introduced into the control law design to remove the negative effect of the input limitation theoretically. In order to improve the performance of the sliding phase, the terminal attractor is incorporated into the design of the ideal sliding surface. We perform numerical simulations to illustrate that the stabilization of the deployment is accomplished by the control law designed in spite of the input limitation.\par
The paper is organized as follows: Section~\ref{sec:2} introduces the standard mathematical model of the deployment of the TSS. Section~\ref{sec:3} designs the adaptive saturated sliding mode control laws with input limitation for the dynamics, and these simulations are exhibited in Section~\ref{sec:4} to reveal the effectiveness and efficiency. The conclusion is given in Section~\ref{sec:5}.
\section{Mathematical model}\label{sec:2}
Figure \ref{fig:1} shows all the elements of the TSS via the geometric representation. In the figure, the axes of the coordinate frame $OXYZ$ passes through the center of the earth $O$. $OX$ and $OY$ are mutually perpendicular axes on the equatorial plane with $OX$ pointing to the spring equinox, and $OZ$ is perpendicular to the equatorial plane, bundling the axis of the earth upwards. The axis $OY$ completes the right-hand triad. The orbital coordinate frame is denoted as $O_2xyz$ with the origin $O_2$ located at the mass center of the TSS. $O_2x$ is directed in the direction of orbital velocity, meanwhile $O_2z$ is defined to point towards the center of the earth, generating the axis $O_2y$ by utilizing the right-hand triad. The origin of the TSS coordinate frame $O_2x'y'z'$ is also fixed at the mass center of the TSS, and the axis $O_2z'$ is directed along the tether and points to the subsatellite. The orientation of $O_2xyz$ relative to the coordinate frame $O_2x'y'z'$ indicates the in-plane and out-of-plane dynamics, where $\theta$ and $\phi$ stand for the in-plane angle and out-of-plane angle, respectively. $\bm R_0$ denotes the position vector between $O_2$ and $O$.

In the analysis of the deployment of the TSS, there is a primary assumption that the movement of entire system is similar to a ``rolling ball'' behavior with three DOFs modelled, and it is reasonable to invoke the spherical coordinates to describe the behavior~\cite{Williams2008}. During space mission, the mass of the tether is much smaller than that of the mother satellite, and hence it is always ignored sufficiently. Some other assumptions are obvious and as follows. The sizes of satellites are small compared with the length of the tether, and hence a satellite can be regard as a particle if the tether is long enough. Actually, the mass ratio between the subsatellite and the mother satellite is tiny, i.e., in YES2, the mother satellite mass is 6530~$kg$, and the subsatellite mass is 12~$kg$~\cite{Williams2008}.  Owing to the tiny mass ratio between the satellites, the deployment of the subsatellite will not affect the motion of the mother satellite during the whole mission. The deployment is regulated by the tension and thrusters, and to guarantee the establishment of the ``rolling ball'' model, the tether remains inextensible, straight and stiff all the time.\par
Symbols used in the model are defined as follows. $R_0$ is a positive scaler meaning the radius of orbit, and $\Omega = \sqrt{\mu_e/R_0^3}$ is the orbital angular velocity where $\mu_e$ is the coefficient of the earth gravity field.\par
%Based on the assumption above, the kinetic energy and the potential energy of the deployment dynamics can be described as~\cite{Mantellato2015}:
\begin{figure}[http]
\centering
\includegraphics[width=0.7\textwidth]{paper4_fig1.eps}
\caption{TSS geometric representation}
\label{fig:1}
\end{figure}
%\begin{align}
%T_t
%&= \frac{1}{2}\sum\limits_{s=1}^2 m_s\bm{R_s'\cdot R_s'}\notag\\
%&= \frac{1}{2}m\bm{R_0'\cdot R_0'}+\frac{1}{2}\sum\limits_{s=1}^2 m_s\bm{r_s'\cdot r_s'}\notag\\
%&= \frac{1}{2}m\Omega^2R_0^2+\frac{1}{2}\bar{m}l^2\begin{bmatrix}\phi'+(\theta'+\Omega)^2\end{bmatrix}+\frac{1}{2}\bar{m}l'^2\label{eq:Tt1}\\
%V_g
%&= -\mu_e\sum\limits_{s=1}^2\frac{m_s}{\vert \bm{R_0+r_s}\vert}\notag\\
%&= -m\Omega^2R_0^2+\frac{\bar{m}\Omega^2l^2}{2}(1-3\cos{^2\theta}\cos^2\phi),\label{eq:Vg}
%\end{align}
%where, $m = m_1+m_2$ stands for the total mass of the system, $m_1$ and $m_2$ of which denote the mass of the mother satellite and of the subsatellite, respectively. $(\cdot)'$ represents the differentiation with respective to time, and $l$ means the tether length connecting the satellites. $\theta$ and $\phi$ depict the rolling and pitching dynamics of the TSS. For clarity, $m_1m_2/m$ is rewritten as $\bar m$.
%Combining Eq.~(\ref{eq:Vg}) and Eq.~(\ref{eq:Tt1}) give rise to the Lagrange function~\cite{Mantellato2015}:
%\begin{align}
%\begin{split}
%L=&T_t-V_g\\
%=&\frac{\bar ml^2}{2}\begin{pmatrix}\phi'^{2}+\cos^2\phi(\theta'+\Omega)^2\end{pmatrix}+\frac{\bar ml'^2}{2}\\
%&-\frac{\bar m\Omega^2l^2}{2}\begin{pmatrix}1-3\cos^2\theta\cos^2\phi\end{pmatrix}+\frac{3m\Omega^2R_0^2}{2},
%\end{split}
%\end{align}
Based on the assumption mentioned above and Lagrangian mechanics theory, we can directly obtain the following expression for the TSS~\cite{Sun2014304,Mantellato2015}
\begin{align}
\begin{split}
&\bar m l''-\bar m l\left\{\begin{bmatrix}\phi'^2+(\theta'+\Omega)^2\cos^2\phi\end{bmatrix}+\Omega^2(3\cos^2\phi\cos^2\theta-1)\right\}=-\tau_t+\tau_{td}\\
&\bar m l^2\cos^2\phi\theta''+2\bar m (\theta'+\Omega)l^2\cos^2\phi(l'/l-\phi'\tan\phi)+3\bar m \Omega^2l^2\sin\theta\cos\theta\cos^2\phi=\tau_\theta+\tau_{\theta d}\\
&\bar m l^2\phi''+2\bar m \phi'l'l+\bar m l^2\sin\phi\cos\phi\left[(\theta'+\Omega)^2+3\bar m \Omega\cos^2\theta\right]=\tau_\phi+\tau_{\phi d},\label{eq:dynamics 1}
\end{split}
\end{align}
where, $m = m_1+m_2$ stands for the total mass of the system, $m_1$ and $m_2$ of which denote the mass of the mother satellite and of the subsatellite, respectively. $(\cdot)'$ represents a first derivative with respect to time, and $l$ means the tether length connecting the satellites. $\theta$ and $\phi$ depict the roll and pitch dynamics of the TSS. For clarity, $m_1m_2/m$ is rewritten as $\bar m$. $\tau_t$, $\tau_\theta$ and $\tau_\phi$ are the tension, the in-plane thruster force and the out-of-plane thruster force, respectively. $\tau_{t d}$, $\tau_{\theta d}$ and $\tau_{\phi d}$ stand for the bounded external disturbances correspondingly. Typically, the dimensionless transformations simplify issues for researchers to design the deployment scheme, and the conventional transformations are represented as shown below~\cite{williams2009yes2,wen2015space,wen2016constrained}:
\begin{align}
\begin{split}
&\lambda=l/L\\
&\rm{d}\it{()}/\rm{d}\it{t}=\Omega \rm{d}\it{()}/\rm{d}\it{v}\\
&\hat{\tau}_{t0}=-\tau_t/(\bar{m}\Omega^2L)\\
&\hat{\tau}_\theta=\tau_\theta/(\bar{m}\Omega^2L^2)\\
&\hat{\tau}_\phi = \tau_\phi/(\bar{m}\Omega^2L^2)\\
&\hat{\tau}_{td}=\tau_{td}/(\bar{m}\Omega^2L)\\
&\hat{\tau}_{\theta d}=\tau_{\theta d}/(\bar{m}\Omega^2L^2)\\
&\hat{\tau}_{\phi d}= \tau_{\phi d}/(\bar{m}\Omega^2L^2)\label{eq:varcon0}.
\end{split}
\end{align}
From~(\ref{eq:varcon0}), it is easy to find that the boundedness of $\hat{\tau}_{t0}$ is different from $\hat{\tau}_\theta$'s and $\hat{\tau}_\phi$'s, since $\hat{\tau}_{t0}$ just owns the negative boundedness due to positive $\tau_t$. However, this issue doesn't conform to conventional hypotheses about input saturation, i.e., for adaptive sliding mode control law design, $\hat{\tau}_{t0}$ doesn't fulfil the requirement of transformation of limited inputs, leading to the nonrigorous stability analysis~\cite{Hu2008552,6060930}. To cope better with the missing positive boundedness, we initiatively introduce the dimensionless length $\lambda$ into the dimensionless tension to establish a reasonable positive boundedness, and hence limited generalized inputs are produced by the novel dimensionless transformations:
\begin{align}
\begin{split}
&\lambda=l/L\\
&\rm{d}\it{()}/\rm{d}\it{t}=\Omega \rm{d}\it{()}/\rm{d}\it{v}\\
&\hat{\tau}_t=-\tau_t/(\bar{m}\Omega^2L)+\lambda\\
&\hat{\tau}_\theta=\tau_\theta/(\bar{m}\Omega^2L^2)\\
&\hat{\tau}_\phi = \tau_\phi/(\bar{m}\Omega^2L^2)\\
&\hat{\tau}_{td}=\tau_{td}/(\bar{m}\Omega^2L)\\
&\hat{\tau}_{\theta d}=\tau_{\theta d}/(\bar{m}\Omega^2L^2)\\
&\hat{\tau}_{\phi d}= \tau_{\phi d}/(\bar{m}\Omega^2L^2)\label{eq:varcon}.
\end{split}
\end{align}
The obvious benefits of these transformations are that all the inputs own the positive and negative boundedness and hence satisfy the assumption about limited inputs of existing control laws~\cite{Hu2008552,6060930}. Furthermore, the generalized inputs~(\ref{eq:varcon}) can be processed by the proposed Sine function in this paper to yield the strict saturated inputs, which essentially guarantee the rigorous stability analysis of the proposed adaptive sliding mode control law.

Based on the novel dimensionless transformations~(\ref{eq:varcon}), Eq.~(\ref{eq:dynamics 1}) can be converted into the dimensionless Euler-Lagrange dynamical equation described as follows:
\begin{align}
H(\bm q)\ddot {\bm q}+C(\bm q,\dot{\bm q})\dot{\bm q}+G(\bm q) = \bm\tau+\bm\tau_d,\label{eq:dynamic model}
\end{align}
where
\begin{align}
\begin{split}
\bm q&= \begin{pmatrix}\lambda&\theta&\phi\end{pmatrix}^T\\
\bm \tau&=\begin{pmatrix}\hat\tau_t&\hat\tau_\theta&\hat\tau_\phi\end{pmatrix}^T\\
\bm\tau_d&=\begin{pmatrix}\hat\tau_{td}&\hat\tau_{\theta d}&\hat\tau_{\phi d}\end{pmatrix}^T\\
H(\bm q) &= \begin{bmatrix}1 &0&0\\0 &\lambda^2\cos^2\phi&0\\0&0&\lambda^2\end{bmatrix}\\
C(\bm q,\dot{\bm q}) &=\begin{bmatrix}0 &-(\lambda\dot \theta+2\lambda)\cos^2\phi&-\lambda\dot\phi\\(2\lambda+\lambda\dot \theta)\cos^2\phi&\lambda\dot \lambda\cos^2\phi-\lambda^2\dot\phi\sin\phi\cos\phi&-(\dot\theta+2)\lambda^2\sin\phi\cos\phi\\ \lambda\dot\phi&\lambda^2(\dot\theta+2)\sin\phi\cos\phi&\lambda\dot \lambda\end{bmatrix}\\
G(\bm q) &=\begin{bmatrix}-\lambda\cos^2\phi+2\lambda+3\lambda\cos ^2\theta\cos^2\phi\\3\lambda^2\cos\theta\sin\theta\cos^2\phi\\\lambda^2+3\lambda^2\cos^2\theta\sin\phi\cos\phi\end{bmatrix}.
\end{split}
\end{align}
Based on the expression above, one can find that $\dot H(\bm q) - 2C(\bm q,\dot{\bm q})$ is skew-symmetric, namely,
\begin{align}
\bm q^T(\dot H(\bm q) - 2C(\bm q,\dot{\bm q}))\bm q = 0.
\end{align}
\begin{myrem}

For the deployment mission, the following expression always exists,
\begin{align}
0<\lambda_{min}\le\lambda\le \lambda_{max}\label{eq:lambda}.
\end{align}
In reality, once $\lambda = 0$ is satisfied, the mother satellite and subsatellite will collide with each other, which is meaningless for the deployment of the TSS. Hence the minimum of the dimensionless length $\lambda_{min}$ is introduced to maintain the reasonably safe formation of the satellites, and also guarantees the non-singularity of the dimensionless transformation~(\ref{eq:dynamics 1}) theoretically. $\lambda_{max}$ stands for the maximum tether length ratio.\par
The tension acting on the space tether is a positive scalar during the deployment, since zero tension means slack, which is inconsistent with the fundamental assumption that the tether is stiff in the ``rolling ball'' model. $\tau_{tmax}$ denotes the maximum tensile tolerance of the tether, and $\tau_t>\tau_{tmax}$ may lead to the break of the tether. Hence, $\tau_{tmin}$ can be a small value to guarantee the stiff tether, and reasonable $\tau_{tmax}$\ prevents the fracture of the tether.
\end{myrem}
This section has established the deployment dynamics with the input limitation and an adaptive saturated sliding mode control law will be presented in the subsequent section.
\section{Dynamic sliding mode control law subjected to input limitation}\label{sec:3}
Before declaring the control law design, some basic preconditions are prepared for proceeding derivations.  The left side of Eq.~(\ref{eq:dynamic model}) can be denoted by the linearly parametrization $Y\Theta$, in which $Y=Y(\bm q,\dot {\bm q},\ddot {\bm q})$ consists of known nonlinear functions. $\bm\Theta$ denotes unknown but constant parameters, and particularly it is a constant vector in this dimensionless form. Reconstruct the parametrization $Y\bm\Theta$ in terms of a reference $\dot{\bm q}_r$~\cite{parra2003dynamic}:
\begin{align}
 Y_r\bm\Theta\equiv H(\bm q)\ddot {\bm q}_r+C(\bm q,\dot {\bm q})\dot {\bm q}_r+G(\bm q)\label{eq:dynamic model 1}
\end{align}
where $Y_r = Y_r(\bm q,\dot{\bm q},\dot{\bm q}_r,\ddot{\bm q}_r)$. Based on results mentioned above, the open-loop error dynamics expressed by $\bm S_r$ can be obtained, as follows:
\begin{align}
H(\bm q)\dot {\bm S}_r+C(\bm q,\dot{\bm q})\bm S_r=\bm\tau +\bm\tau_d- Y_r\bm\Theta\label{eq:dynamic model 2}
\end{align}
where $\bm S_r$ is defined as
\begin{align}
\bm S_r=\dot {\bm q}-\dot {\bm q}_r.\label{eq:Sr1}
\end{align}
For the system~(\ref{eq:dynamic model 2}) with the input limitation ignored, the author in \cite{parra2003dynamic} has designed a nonlinear dynamic sliding PID controller with the zero initial condition:
\begin{align}\begin{split}
\dot {\bm q}_r&=\dot{\bm q}_d-\alpha\Delta\bm q-\gamma\bm\sigma\\
\dot{\bm\sigma}&=\mathbf{sign}(\bm S)\\
\bm S&=\Delta\dot{\bm q}+\alpha\Delta\bm q\label{eq:pid controller}
\end{split}\end{align}
where
\begin{align}
\mathbf{sign}(\bm x) = \left(\rm{sign}(\it{x_1}),\rm{sign}(\it{x_2}),\cdots,\rm{sign}(\it{x_n})\right)^T
\end{align}
for a vector $\bm x=\left(x_1,x_2,,\cdots,x_n\right)^T$. $\alpha$ and $\gamma$ are control law parameters which will be discussed in the control law design. Define the tracking error as $\Delta \bm q=\bm q-\bm q_d$, in which $\bm q_d$ is the desired trajectory with $\bm q_d(t)\in C^2$. However, the PID controller~(\ref{eq:pid controller}) can not be directly applied to the system~(\ref{eq:dynamic model 2}) with the input limitation, and in the subsequent section, we will utilize the saturated sliding mode technique to modify the controller~(\ref{eq:pid controller}) for the system~(\ref{eq:dynamic model 2}). For deriving the proposed control scheme, some properties should be given out in advance.
\begin{mypro}\cite{parra2003dynamic}\label{pro:1}
Positive scalars $\Xi_i(i = 0,\ldots,6)$ always exist such that
\begin{align}
\begin{split}
\Vert H(\bm q)\Vert & \ge \zeta_{min}(H(\bm q))>\Xi_0>0\\
\Vert H(\bm q)\Vert & \le \zeta_{max}(H(\bm q))<\Xi_1<\infty\\
\Vert C(\bm q,\dot {\bm q})\Vert & \le \Xi_2\Vert\dot {\bm q}\Vert+\Xi_3\\
\Vert G(\bm q)\Vert & \le \Xi_4\\
\Vert \dot {\bm q}_r\Vert & \le \Xi_5+\alpha\Vert\Delta \bm q\Vert+\gamma\Vert\bm\sigma\Vert\\
\Vert \ddot {\bm q}_r\Vert & \le \Xi_6+\alpha\Vert\Delta \dot {\bm q}\Vert\\
\end{split}
\end{align}
where $\zeta_{min}(\cdot)$ and $\zeta_{max}(\cdot)$ mean the minimum and maximum eigenvalues of matrix, respectively. Furthermore, one can have $Y_r\bm\Theta\le \eta(t)$, in which $\eta(t)=f(\Delta \bm q,\Delta \dot {\bm q},\sigma,\Xi_i,t)$ is a state-dependent function, thus $Y_r\bm\Theta$ is usually thought to be a bounded function with $\vert\eta\vert\le\varphi_0$, where $\varphi_0$ is a positive scalar determined according to the system states.
\end{mypro}

\begin{myas}
In low-altitude orbit , the external disturbance $\bm\tau_d$ contains the gravity gradient torque, the aerodynamic interference and the solar radiation pressure torque. Thanks to existing studies, the values of the solar radiation pressure and the aerodynamic torque are about $1\times10^{-5}$ Nm, and meanwhile the gradient torque is in the range of $1\times10^{-5}$ Nm. The space perturbation can be specifically regarded as the periodic function in this paper, and the period equals the orbital period $5000$ s~\cite{Inamori2015192}. Hence, the external disturbance can be described by the following expression
\begin{align}
\bm\tau_d=\left(0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000)\right)^T.
\end{align}
Therefore, the bounded external disturbance satisfies
\begin{align}
\Vert\bm\tau_d\Vert\le\varphi_1<\infty,
\end{align}
where $\varphi_1$ can be selected as a positive scalar $\varphi_1=1$.
\end{myas}
\subsection{Saturated sliding mode control law subjected to input limitation}
The process above can frame a conventional dynamic sliding PID controller, which can regulate the dynamics of n-link robotic manipulator if the dimension in the above derivation  is extended~\cite{parra2003dynamic}. However, it is not possible to regulate the dynamics of the deployment as desired theoretically, because the input $\tau$ in the PID controller is not confined to the characteristics of the tension. To cope with the issue about bounded inputs, we develop the dynamic sliding mode control law with the experience of the saturated manifold, which is also a popular technique in the control of Lagrangian system.\par
Before giving out main results, some routine lemmas emerge in advance to support proceeding derivations.
\begin{mylem}\label{lemma:1}
Consider a quasi-saturated function as follows:
\begin{align}
\rm{Sine}\it{(x)} =
\begin{cases}
\rm{sign}\it{(x)}&\vert x\vert\ge\pi/2\\
\sin(x)&\vert x\vert<\pi/2
\end{cases},
\end{align}
and it's notable that the function above is saturated and owns an important property: $\vert \rm{Sine}\it{(x)}\vert\le\vert x\vert$, $\forall x$. The geometric illustration is shown in Figure~\ref{fig:2}, from which an interesting property can be found by sliding the curve $y=\rm{Sine}(\it{x})$ along the axis $x=0$ in the range $[-1,1]$, and once the function $y=\rm{Sine}\it{(x)}+\Delta h$ with $\vert\Delta h\vert<1$ settles, one can choose the function $y = x+\arcsin(\Delta h)$ to obtain following relations:
\begin{align}
&\vert \rm{Sine}\it{(x)}+\Delta h\vert\le\vert x+\arcsin(\Delta h)\vert\label{eq:lem11}\\
&\rm{sign}(\rm{Sine}\it{(x)}+\Delta h)=\rm{sign}(\it{x}+\arcsin(\Delta h)), \forall x, \vert\Delta h\vert<1\label{eq:lem12}.
\end{align}
\end{mylem}
\begin{proof}
For relation~(\ref{eq:lem11}), let $y=\rm{Sine}\it{(x)}+\Delta h-x-\arcsin(\Delta h)$, and we can get its derivative
\begin{align}
\begin{cases}
\rm{d}\it{y}/\rm{d}\it{x} = -1,&x\ge\pi/2,x\le-\pi/2\\
\rm{d}\it{y}/\rm{d}\it{x} = \cos x - 1,&\vert x\vert<\pi/2,
\end{cases}
\end{align}
where $\rm{d}\it{y}/\rm{d}\it{x}<0$ means $y$ is a monotone decreasing function with respect to $x$. As $x=-\arcsin(\Delta h)$, $y=0$, according to the definition of $\rm{Sine}\it{(x)}$, we proceed to get
$$\left.
\begin{aligned}
x\ge-\arcsin(\Delta h), y\le0&\Rightarrow 0\le \rm{Sine}\it{(x)}+\Delta h\le x+\arcsin(\Delta h)\\
x<-\arcsin(\Delta h), y>0&\Rightarrow 0> \rm{Sine}\it{(x)}+\Delta h>x+\arcsin(\Delta h)\end{aligned}\right\}\\\\
\Rightarrow\vert \rm{Sine}\it{(x)}+\Delta h\vert\le\vert x+\arcsin(\Delta h)\vert.
$$
For Eq.~(\ref{eq:lem12}), considering the case $\rm{Sine}\it{(x)}+\Delta h>0$, when $\vert x\vert<\pi/2$, $\rm{Sine}\it{(x)}+\Delta h>0$ becomes
\begin{align}
\sin(x)+\Delta h>0&\Rightarrow x>-\arcsin(\Delta h)\\
&\Rightarrow x+\arcsin(\Delta h)>0,
\end{align}
since $y=\arcsin(x)$ is a monotone increasing function. We only need to consider $x\ge\pi/2$ instead of $\vert x\vert\ge\pi/2$ for $x\le-\pi/2$ is meaningless in the case $\rm{Sine}\it{(x)}+\Delta h>0$. It is obvious that $x+\arcsin(\Delta h)>0$ due to $\vert \arcsin(\Delta h)\vert<\pi/2$. Therefore, we can get
\begin{align}
\rm{sign}(\rm{Sine}\it{(x)}+\Delta h)=\rm{sign}\it{(x}+\arcsin(\Delta h)),\rm{Sine}\it{(x)}+\Delta h>0,
\end{align}
and similarly the following equation is reasonable:
\begin{align}
\rm{sign}(\rm{Sine}\it{(x)}+\Delta h)=\rm{sign}\it{(x}+\arcsin(\Delta h)),\rm{Sine}\it{(x)}+\Delta h\le0.
\end{align}
\end{proof}
\begin{myrem}Moreover, we think that a quantity $y$ is saturated iff it is produced directly by the saturated function $y=\rm{Sine}\it{(x)}$. Otherwise, the bounded quantity is different from the saturated one since it does not own the standard formulation like $y=\rm{Sine}\it{(x)}$, i.e., a quantity can just be referred to as a bounded quantity if its boundary is $[a,b]$ with $a\neq -1$ or $b\neq 1$.
\end{myrem}
\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{paper4_fig2.eps}
\caption{$y =\rm{Sine}\it{(x)}$ and $y=\rm{Sine}\it{(x)}+\Delta h$}
\label{fig:2}
\end{figure}
\begin{mylem}\cite{Hu2008552}\label{lemma:2}
Consider a bounded function owning the subsequent property:
\begin{align}
g(y) &= R(y)y\in[y_{min},y_{max}]\\
R(y)&=\begin{cases}
y_{max}/y&y>y_{max}\\
1&y_{min}\le y\le y_{max}\\
y_{min}/y&y<y_{min},
\end{cases}
\end{align}
where $y_{min}$ and $y_{max}$ stand for the minimum and maximum values of $g(y)$, respectively, and $y_{min}<0<y_{max}$. If $y$ is bounded, $^by_{min}\le y\le{^by_{max}}$, hence
\begin{align}
0<\min\{y_{min}/{^by_{min}},y_{max}/{^by_{max}}\}\le R(y)\le 1.
\end{align}
\end{mylem}
\begin{myrem}
As we know, the tension, in-plane and out-of-plane thruster forces are all limited during the deployment mission, and we describe these inputs by using the matrix and vector to predigest the following illation.
\begin{align}
\bm{g}(\bm\tau) &= \left(g(\hat\tau_t),g(\hat\tau_\theta),g(\hat\tau_\phi)\right)^T\\
R_M(\bm\tau) &= \begin{bmatrix}R(\hat\tau_t)&0&0\\0&R(\hat\tau_\theta)&0\\0&0&R(\hat\tau_\phi)\end{bmatrix}.
\end{align}
and in order to design the saturated sliding mode control law for the dynamics of the deployment, the $\rm{Sine}$ function for vectors is specified as:
\begin{align}
\mathbf{Sine}(\bm x)=\begin{pmatrix}\rm{Sine}\it{(x_1)},\rm{Sine}\it{(x_2)},\cdots,\rm{Sine}\it{(x_n)}\end{pmatrix}^T.
\end{align}
\end{myrem}
The above lemmas introduce some properties about saturated functions, with which the dynamic sliding mode control law subjected to the input limitation for Lagrangian system~(\ref{eq:dynamic model}) can be designed by using the subsequent theorem.
\begin{mythm}\label{thm:1}(Saturated sliding mode control law)
Considering a nominal Lagrangian system~(\ref{eq:dynamic model}) expressing the dynamics of the deployment of the TSS, one can construct a saturated sliding mode control law subjected to the input limitation, as follows:
\begin{align}
\bm\tau = -\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1})),\label{eq:slider control}
\end{align}
where
\begin{align}
\begin{split}
\bm S_{r1} &= \bm S_{q1}+\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) d\varsigma\\
\bm S_{q1} &= \mathbf{Sine}(\Delta \dot {\bm q}+\alpha\Delta\bm q),
\end{split}
\end{align}
in which, $F(0)>0$, and the derivative of $F(t)$ has the characteristics:
\begin{align}
\frac{\rm{d}}{\rm{d}\it{t}}F(t) = \delta\varphi \bm S_{r2}^T K_1\rm{e}\it{^{-2K_1^{-1}F^{-1}(t)}}\mathbf{sign}(\bm S_{r2})
\end{align}
where $K_1\in R^{3\times 3}$ is a positive definite diagonal matrix, which will be given out later, and if $K\in R^{3\times 3}$, define
\begin{align}
\rm{e}\it{^K}=
\begin{bmatrix}
\rm{e}\it{^{K_{11}}}&\rm{e}\it{^{K_{12}}}&\rm{e}\it{^{K_{13}}}\\
\rm{e}\it{^{K_{21}}}&\rm{e}\it{^{K_{22}}}&\rm{e}\it{^{K_{23}}}\\
\rm{e}\it{^{K_{31}}}&\rm{e}\it{^{K_{32}}}&\rm{e}\it{^{K_{33}}}
\end{bmatrix}.
\end{align}
$\varphi$ is the boundedness of the lumped uncertainty. $\gamma$ is designed as a time-varying positive scalar and $\delta>1$ is the design parameter, so that the deployment dynamics can be regulated as desired.
\end{mythm}
\begin{proof}
Consider the dynamic coordinate change $\bm S_{r2}$ with regular error vectors $\Delta \bm q$ and $\Delta\dot{\bm q}$:
\begin{align}
\begin{split}
\bm S_{r2} &= \bm S_{q2}+\bm{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma})\\
\bm S_{q2} &= \Delta \dot {\bm q}+\alpha \Delta \bm q,
\end{split}
\end{align}
where $\bm y=\mathbf{arcsin}(\bm x)$ is the vector function with
\begin{align}
\mathbf{arcsin}(\bm x) = \left(\arcsin(x_1),\arcsin(x_2),\cdots,\arcsin(x_n)\right)^T
\end{align}
for a vector $\bm x=\left(x_1,x_2,\cdots,x_n\right)^T$. It is obvious that $\bm S_{q1}$ is exactly saturated, hence $\bm S_{r1}$ is also bounded if the dynamics can be controlled as desired. In addition, for guaranteeing that $\bm x=\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) d\varsigma$ always falls in the field of definition of $\bm y=\mathbf{arcsin}(\bm x)$, $\gamma$ can be designed as a time-varying parameter
    \begin{align}
    \gamma=\begin{cases}
    \gamma_0&\Vert\int^t_{t_0}\bm S_{q1}(\varsigma) d\varsigma\Vert\le 1\\
    \gamma_0\Vert\int^t_{t_0}\bm S_{q1}(\varsigma) d\varsigma\Vert^{-1}&\Vert\int^t_{t_0}\bm S_{q1}(\varsigma) d\varsigma\Vert> 1,
    \end{cases}
    \end{align}
    where $\gamma_0$ can be selected as $\gamma_0=0.1$.

According to Lemma~\ref{lemma:1}, it is apparent that $\bm S_{r2}$ and $\bm S_{r1}$ have the subsequent relations:
\begin{align}
&\bm S_{r2}^T \mathbf{sign}(\bm S_{r2}) \ge  \bm S_{r1}^T \mathbf{sign}(\bm S_{r1})\\
&\mathbf{sign}(\bm S_{r2})=\mathbf{sign}(\bm S_{r1}),
\end{align}
Replacing the dynamic coordinate change $\bm S_r$ in Eq.~(\ref{eq:dynamic model 2}) with $\bm S_{r2}$ results in the following expression:
\begin{align}
H(q)\dot {\bm S}_{r2} = -\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1}))+\bm\tau_d-C(q,\dot q)\bm S_{r2}-Y_{r2}\bm\Theta.
\end{align}
Similar to Properties~\ref{pro:1}, one can find a positive scalar $\varphi_2$ such that $\Vert Y_{r2}\bm\Theta\Vert\le\varphi_2$, and moreover the boundedness of the lumped uncertainty $\varphi$ satisfies $\varphi>\varphi_1+\varphi_2$.

Select the Lyapunov function candidate
\begin{align}
V=\frac{1}{2}\bm S_{r2}^TH(q)\bm S_{r2}+\frac{1}{2}v^2,
\end{align}
where $v=F(t)\rm{e}\it{^{K_1^{-1}F^{-1}(t)}}$. Differentiating the Lyapunov function above gives rise to
\begin{align}
\begin{split}
\dot V &= \bm S_{r2}^TH(q)\dot {\bm S}_{r2}+\frac{1}{2}\bm S_{r2}^T\dot H(q)\bm S_{r2}+v\dot v\\
&= -\bm S_{r2}^T\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1}))+\bm S_{r2}^T\bm\tau_d-\bm S_{r2}^TC(q,\dot q)\bm S_{r2}+\frac{1}{2}\bm S_{r2}^T\dot H(q)\bm S_{r2}-\bm S_{r2}^TY_{r2}\bm\Theta+v\dot v\\
&= -\bm S_{r2}^T\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1}))+\bm S_{r2}^T\bm\tau_d-\bm S_{r2}^TY_{r2}\bm\Theta+\frac{1}{2}\bm S_{r2}^T(\dot H(q)-2C(q,\dot q))\bm S_{r2}+v\dot v\\
&= -\bm S_{r2}^T\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1}))+\bm S_{r2}^T\bm\tau_d-\bm S_{r2}^TY_{r2}\bm\Theta+v\dot v\\
&= -\bm S_{r2}^TR_M(\delta\varphi F(t) \mathbf{sign}(\bm S_{r1}))\delta\varphi F(t) \mathbf{sign}(\bm S_{r1})+\bm S_{r2}^T\bm\tau_d-\bm S_{r2}^TY_{\bm r2}\bm\Theta+v\dot v.\label{eq:V1_1}
\end{split}
\end{align}
Since the command signal $ \bm\tau_c=-\delta\varphi F(t) \mathbf {sign}(\bm{S}_{r1})$ is bounded, we can select reasonably small positive scalars $c_1$, $c_2$ and $c_3$ which satisfy $0<c_1\le R(\hat\tau_t)<1$, $0<c_2\le R(\hat\tau_\theta)<1$ and $0<c_3\le R(\hat\tau_\phi)<1$, and obtain the diagonal matrix $K_1$ as follows:
\begin{align}
K_1 = \begin{bmatrix}c_1&0&0\\0&c_2&0\\0&0&c_3\end{bmatrix}.\label{eq:saturated ratio}
\end{align}
$v\dot v$ can be derived as follows:
\begin{align}
v\dot v =\varphi\delta \bm S_{r2}^T K_1(F(t)-K_1^{-1})\mathbf{sign}(\bm S_{r2}),
\end{align}
considering $\mathbf{sign}(\bm S_{r1})=\mathbf{sign}(\bm S_{r2})$, we can proceed to rewrite Eq.~(\ref{eq:V1_1}) as:
\begin{align}
\begin{split}\label{eq:V1_2}
\dot V&\le -\bm S_{r2}^T\delta\varphi  K_1F(t)\mathbf{sign}(\bm S_{r2})+\bm S_{r2}^T(\varphi_1+\varphi_2) \mathbf{sign}(\bm S_{r2})+\varphi\delta \bm S_{r2}^T K_1(F(t)-K_1^{-1})\mathbf{sign}(\bm S_{r2})\\
&\le \varphi \bm S_{r2}^T\mathbf{sign}(\bm S_{r2})-\varphi\delta \bm S_{r2}^T \mathbf{sign}(\bm S_{r2})\\
&\le\varphi(1-\delta)\bm S_{r2}^T\mathbf{sign}(\bm S_{r2})\le 0
\end{split}
\end{align}
where a large enough $\delta>1$ can be selected to satisfy $\dot V \le 0$. According to the description of Barbalat's lemmas, we can conclude  that $\bm S_{r2}$ asymptotically converges to the origin. Also, the finite reaching time can be calculated by referring to~\cite{Ma201667}.

Utilizing $\bm S_{r2}=0$ leads to the equation
\begin{align}
\bm S_{q2}=-\mathbf{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma}),\label{eq:eqSq2}
\end{align}
and for employing Lyapunov second method to analyse the stabilities of $\bm S_{q2}$ and $\Delta\bm q$, we firstly select a Lyapunov candidate function
\begin{align}
V_1 = \frac{1}{2}\bm S_{q2}^T\bm S_{q2},
\end{align}
and the derivative of which gives rise to
\begin{align}
\begin{split}
\dot V_1 &= \bm S_{q2}^T\dot {\bm S}_{q2}\\
&=-\frac{\gamma \bm S_{q2}^T\bm S_{q1}}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}\\
&\le -\frac{\gamma \Vert \bm S_{q1}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}\le 0,
\end{split}
\end{align}
if and only if $\bm S_{q1} = 0$, accordingly, $\bm S_{q2} = 0$, one has $\dot V_1=0$. Hence the asymptotically stability of Eq.~(\ref{eq:eqSq2}) has been proved, and therefore $\bm S_{q2}\rightarrow 0$ as $t\rightarrow \infty$ induces
\begin{align}
\Delta \dot {\bm q}=-\alpha \Delta \bm q.
\end{align}
We proceed to select the Lyapunov function candidate $V_2=\frac{1}{2}\Delta \bm q^T\Delta\bm q$ for deriving the asymptotically stability by using Lyapunov second method, and the derivative of $V_2$ follows
\begin{align}
\begin{split}
\dot V_2 &= \Delta\bm q^T\Delta \dot{\bm q}\\
&=-\alpha \Delta \bm q^T\Delta \bm q\le 0,
\end{split}
\end{align}
if and only if $\Delta \bm q = 0$, namely, $\Delta \dot {\bm q} = 0$, one has $\dot V_2=0$. Hence asymptotically convergence of $\Delta \bm q$ has been explained, namely, $\Delta \bm q\rightarrow 0$ as $t\rightarrow \infty$. To conclude, the error state $\Delta \bm q$ of the system~(\ref{eq:dynamics 1}) asymptotically converges to the origin with the limited input~(\ref{eq:slider control}). This completes the proof.
\end{proof}
\begin{myrem}
For the adaptive sliding mode control law with input limitation, the adaptive update rate $\frac{\rm{d}}{\rm{d}\it{t}}F(t)$ always helps to remove the side-effect of the input limitation during the stability analysis by using Lyapunov second method, and moreover, the actual examples of $F(t)$ are unhelpful to the theoretical derivation and the control law design~\cite{Hu2008552,6060930}. This is because we can search the actual value of $F(t)$  by the way of state updating from the initial condition $F(0)=0$, for instance, in our simulations, we apply Ode45 technique to update states of the overall system including this adaptive update rate.
\end{myrem}
In the aforementioned controller design, signum functions in inputs lead to chattering phenomena, which may cause adverse unmodeled dynamics to ruin the control system. Herein, we adopt the boundary layer technique to reduce chattering through replacing $\rm{sign}\it{(x)}$ by $\rm{sat}\it{(x/\nu)}$, where
\begin{align}
\rm{sat}\it{(x/\nu)}=
\begin{cases}
x,&\vert x\vert <\nu\\
\rm{sign}\it{(x)},&\vert x\vert\ge\nu
\end{cases}
\end{align}
and $\nu$ is the boundary layer thickness. However, the cost of using boundary layer in sliding mode controller is that the asymptotic stability of proposed method can not be guaranteed, and we proceed to analyze the convergence precision in the following theorem.
\begin{mythm}
Considering the controller in Theorem~\ref{thm:1}, we replace $\mathbf{sign}{(\bm{x})}$ in Eq.~(\ref{eq:slider control}) by $\mathbf{sat}{(\boldsymbol {x/\nu})}$, where $\mathbf{sat}{(\boldsymbol{x}/\boldsymbol{\nu})}=\left(\rm{sat}\it{(x_1/\nu_1)},\rm{sat}\it{(x_2/\nu_2)},\cdots,\rm{sat}\it{(x_n/\nu_n)}\right)^T$, to remove the chattering caused by the signum function, and the modified controller can guarantee the precisions of system states $\Delta\bm q$ that satisfy
\begin{align}
  \Vert\Delta\bm q\Vert\le\frac{\Vert\boldsymbol{arcsin}(\bm \nu)\Vert}{\alpha}.
\end{align}
$\bm z$ is the position vector that describes the distances of system states in the boundary layer from the sliding mode surface, and $\dot{\bm z}$ is the velocity vector correspondingly. It is obvious that $\bm z$ is bounded and has $\Vert\bm z\Vert\le\Vert\bm\nu\Vert$, where $\bm\nu=(\nu_1,\nu_2,\nu_3)^T,\nu_1=\nu_2=\nu_3\ll 1$ is a constant vector denoting the boundary layer thicknesses for the three channels. Since all the inputs are bounded, it is evident to assume that $\dot{\bm z}$ is bounded with $\Vert\dot{\bm z}\Vert\le \bar z\le 2\Vert\bm\nu\Vert/\rm{d}\it{t}$.
\end{mythm}
\begin{proof}
Referring to the derivation in Theorem~\ref{thm:1}, we can guarantee the system states reach and enter the boundary layer $\Vert\bm{ S}_{r2}\Vert \le \Vert\bm\nu\Vert$ such that
\begin{align}
\bm S_{q2}=\bm z-\mathbf{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma}).
\end{align}
The selected Lyapunov function $V_3=\frac{1}{2}\bm{ S}_{q2}^T\bm{ S}_{q2}$ gives rise to
\begin{align}\begin{split}
\dot V_3&=\bm{ S}_{q2}^T\dot{\bm{S}}_{q2}\\
&=-\frac{\gamma \bm S_{q2}^T\bm S_{q1}}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\bm{ S}_{q2}^T\dot{\bm z}\\
&\le -\frac{\gamma \Vert \bm S_{q1}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\Vert\dot{\bm{z}}\Vert(\Vert\bm z\Vert+\Vert\mathbf{arcsin}(\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma})\Vert)\\
&\le -\frac{\gamma \Vert \bm S_{q1}\Vert^2}{\sqrt{1-\Vert\gamma\int^t_{t_0}\bm S_{q1}(\varsigma) \rm{d}\it{\varsigma\Vert^2}}}+\bar z(\Vert\bm{\nu}\Vert+\pi/2)\\
\end{split}\end{align}
Based on the Lyapunove stability theory, the above expression implies that $\bm S_{q2} = \Delta \dot {\bm q}+\alpha \Delta \bm q=0$ can attract the system states under the condition
\begin{align}
  \Vert \bm S_{q1}\Vert\ge\sqrt{\frac{\bar z(\Vert\bm{\nu}\Vert+\pi/2)}{\gamma}}
\end{align} Rephrasing that, the controller in Theorem~\ref{thm:1} with boundary layer technique can guarantee $\bm S_{q1}$ to converge to $\Vert\bm S_{q1}\Vert\le\sqrt{\frac{\bar z(\Vert\bm{\nu}\Vert+\pi/2)}{\gamma}}$, and once the states reach the boundary layer, since $\bm S_{q1}=\mathbf{Sine}(\Delta\dot{\bm q}+\alpha\Delta\bm q)$ we can combine
\begin{align}
  \mathbf{Sine}(\Delta\dot{\bm q}+\alpha\Delta\bm q)=\bm\nu
\end{align}
with the Lyapunov function $V_4=\frac{1}{2}\Delta\bm q^T\Delta\bm q$ to obtain
\begin{align}\begin{split}
\dot V_4 &= \Delta\bm q\Delta\dot{\bm q}\\
&=\Delta\bm q^T(-\alpha\Delta\bm q+\boldsymbol{arcsin}(\bm \nu))\\
&\le-\alpha\Vert\Delta\bm q\Vert^2+\Vert\Delta\bm q\Vert\Vert\boldsymbol{arcsin}(\bm \nu)\Vert\\
&\le-(\alpha\Vert\Delta\bm q\Vert-\Vert\boldsymbol{arcsin}(\bm \nu)\Vert)\Vert\Delta\bm q\Vert,
\end{split}\end{align}
which indicates that $\Vert\Delta\bm q\Vert=\frac{\Vert\boldsymbol{arcsin}(\bm \nu)\Vert}{\alpha}$ is an attractor, and hence, the ultimately system state precision of the controller with boundary layer technique is
\begin{align}
  \Vert\Delta\bm q\Vert\le\frac{\Vert\boldsymbol{arcsin}(\bm \nu)\Vert}{\alpha}.
\end{align}
This completes the proof.
\end{proof}
\subsection{Quasi-saturated terminal sliding mode control law subjected to input limitation}
The design methodology of the saturated sliding mode control law has been exhibited in last subsection, and the entire strict proof ensures that the deployment dynamics of the TSS can be regulated with the limited tension control. If the order of the integrated error becomes a fraction, the value of which is less than one, we obtain the terminal attractor like
\begin{align}
\dot Z = Z^{p/q},
\end{align}
where $0<p/q<1$ while $p$ and $q$ are both positive odd, and a equivalent formula
\begin{align}
\dot Z = \vert Z\vert^{\hat\alpha} {\rm{sign}}\it{(Z)},
\end{align}
where $0<\hat \alpha<1$, performs the same behavior. Putting the terminal attractor into the saturated sliding mode control law yields a saturated slider-like scheme, and before introducing the input limitation into the system, an assumption is given firstly.
\begin{myas}\label{as:1}
For Eq.~(\ref{eq:dynamic model}), the length rate is thought to be bounded during the deployment, namely, $0<\epsilon\le\dot \lambda<\infty$, such that $\Delta \dot \lambda$ is bounded, either.
\end{myas}
\begin{mythm}\label{thm:2}(Quasi-saturated terminal sliding mode control law)
Considering the deployment dynamics of the TSS as described in Eq.~(\ref{eq:dynamic model}) with limited tension control, the quasi-saturated terminal sliding mode control input can be designed as follow:
\begin{align}
\tau = -\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r3}))
\end{align}
where
\begin{align}
\bm S_{r3} &= \bm S_{q3}+\gamma\int^t_{t_0}\vert \bm S_{q3}(\varsigma)\vert^{\hat\alpha}  \mathbf{sign}(\bm S_{q3}(\varsigma))\rm{d}\it{\varsigma}\label{eq:Sr3}\\
\bm S_{q3} &= \Delta \dot {\bm q}+\alpha \mathbf{Sine}(\xi\Delta \bm q),\label{eq:Sq3}
\end{align}
where $0<\hat\alpha< 1$, such that the dynamics~(\ref{eq:dynamic model 2}) can be regulated as desired. $\bm y=\vert \bm x\vert^{\hat\alpha}$ is defined as $\vert \bm x\vert^{\hat\alpha} =\left(\vert x_1\vert^{\hat\alpha},\vert x_2\vert^{\hat\alpha},\cdots,\vert x_n\vert^{\hat\alpha}\right)^T$. $\varphi$ is the boundedness of the lumped uncertainty, and can be obtained by referring to Theorem~\ref{thm:1}.
\end{mythm}
\begin{proof}
The proof is similar to Theorem~\ref{thm:1}, and omitted here.
\end{proof}
\begin{myrem}
If the design parameter $\hat\alpha = 1$, the quasi-saturated terminal sliding mode control law degenerates into the quasi-saturated sliding mode control law. Compared with the saturated sliding mode control law, $\bm S_{q3}$ converges to the origin in finite time in the quasi-saturated terminal method. This is because the establishment of the equation $\dot {\bm S}_{q3} = \gamma\vert \bm S_{q3}\vert^{\hat\alpha} \mathbf{ sign}(\bm S_{q3})$.
\end{myrem}
\begin{myrem}
Assumption~\ref{as:1} guarantees that the input $\tau$ is bounded in Theorem~\ref{thm:2}, accordingly, a tiny positive scalar $c$ can be found to ensure that $0<c\le R(\hat\tau_t)<1$ always exists. Admittedly, such kind of assumption is skillful, but meaningful to arbitrary controllable dynamics in reality. It is reasonable to propose the terminal sliding mode control law for the dynamics as follows:
\begin{align}
\tau = -\bm{g}(\delta\varphi F(t) \mathbf{sign}(\bm S_{r4})),
\end{align}
where
\begin{align}
\begin{split}
\bm S_{r4} &= \bm S_{q4}+\gamma\int^t_{t_0}\bm S_{q4}(\varsigma) \rm{d}\it{\varsigma}\\
\bm S_{q4} &= \mathbf{Sine}(\Delta \dot {\bm q}+\alpha\xi\vert\Delta \bm q\vert^{\hat\alpha}\mathbf{sign}(\Delta \bm q)),\label{eq:Sr4}
\end{split}
\end{align}
so that the finite convergence of the system~(\ref{eq:Sr4}) can be ensured by using terminal sliding mode theory. $\varphi$ is the boundedness of the lumped uncertainty, and can be obtained by referring to Theorem~\ref{thm:1}.
\end{myrem}
\section{Simulation studies}\label{sec:4}
This section presents the simulations of the regulation of the deployment dynamics with saturated sliding mode control, saturated terminal sliding mode control and quasi-saturated terminal sliding mode control laws, and for clarity, these control laws are abbreviated as S-SMC, ST-SMC and QSF-SMC, respectively. The parameters of these control laws can be selected as follows: $\delta=1.1$, $\varphi= 4$, $\varphi_1 = 1$, $\varphi_2= 2$, $\alpha = 0.8$ and $\hat\alpha=0.6$.\par
Besides the design parameter determined, the simulation initial and desired conditions can be described as follows. The circular orbital altitude is selected around 260 $km$ approximately, and the orbital angular velocity is about $1.17\times 10^{-3}$ $rad/s$. At the initial ejection of the subsatellite, the true anomaly is assumed to be 0 $rad$. Meanwhile, the initial deployment velocity is about 3 $m/s$, and the subsatellite is going to be deployed at 3.5 $km$ vertically. The pitch and roll dynamics are both varying in a small range, and hence, without loss of generality, the numerical initial conditions are designed as $\theta_0 = 0.1$, $\phi_0 = 0.1$ and $\lambda_0 = 0.02$; the desired condition is set to be $\theta_d = 0$, $\phi_d = 0$ and $\lambda_d = 1$, accordingly, $\lambda(t_0)\rightarrow\lambda_d$, $t\rightarrow\infty$ with $\theta(t_0)\rightarrow\theta_d$, $\phi(t_0)\rightarrow\phi_d$, $t\rightarrow\infty$. Hence the dimensionless tether length is varying in the range of $[0.02,1]$.

During the deployment, the dimensionless thruster forces which manage the in-plane angle and the out-of-plane angle are assumed to be saturated with $\vert\hat\tau_{\theta}\vert\le5 $ and $\vert\hat\tau_{\phi}\vert\le 5$. Also, the dimensionless tension $\hat\tau_t$ is limited, and from Eq.~(\ref{eq:varcon}), we can obtain
\begin{align}
\begin{split}
&0<\tau_{tmin}/(\bar{m}\Omega^2L)\le-\hat\tau_t+\lambda\le\tau_{tmax}/(\bar{m}\Omega^2L)\\
\Leftrightarrow
&\begin{cases}
\max\{-\hat\tau_t+\lambda\}\le\tau_{tmax}/(\bar{m}\Omega^2L)\\
\min\{-\hat\tau_t+\lambda\}\ge\tau_{tmin}/(\bar{m}\Omega^2L)
\end{cases}\\
\Leftrightarrow
&\begin{cases}
-\hat\tau_{tmin}+\lambda_{max}\le\tau_{tmax}/(\bar{m}\Omega^2L)\\
-\hat\tau_{tmax}+\lambda_{min}\ge\tau_{tmin}/(\bar{m}\Omega^2L).
\end{cases}
\end{split}
\end{align}
Hence, the limited dimensionless tension gives rise to
\begin{align}
-\tau_{tmax}/(\bar{m}\Omega^2L)+\lambda_{max}\le\hat\tau_{tmin}\le\hat\tau_t\le\hat\tau_{tmax}\le-\tau_{tmin}/(\bar{m}\Omega^2L)+\lambda_{min}.
\end{align}
Based on the expression above, one can obtain $\bar{m}\Omega^2L\approx 4.79\times 10^{-3}$, and suitable boundary values $\hat\tau_{tmin}$ and $\hat\tau_{tmax}$ can be chosen to make the dimensionless tension $\hat\tau_t$ fall in the region $[-3,0.01]$.\par
At this point, the simulations of the deployment dynamics of the TSS can be carried on. We firstly conduct the simulation about the pure deployment dynamics without any uncertainties to verify the effectiveness of the proposed control laws, and the results are illustrated in Figure~\ref{fig:3} to~\ref{fig:6}. Figure~\ref{fig:3} shows the dynamics of the length in different control laws, and it is obvious that all the methods can realize the deployment in aspect of the tether length. ST-SMC and QST-SMC complete the deployment at about 4 $rad$. As shown in Figure~\ref{fig:4}, S-SMC and QST-SMC have bigger length rate than ST-SMC before 1.2 $rad$, indicating the curves of the tether length climb more rapidly at the beginning of the deployment. However, the length rate of ST-SMC doesn't decline dramatically like the other two methods after 1.2 $rad$. This slowly varying length rate yields a smooth trajectory of the deployment shown in Figure~\ref{fig:3}, and meanwhile the length rate converges to the origin earlier than S-SMC and QST-SMC.\par
The in-plane and out-of-plane angles reflect the swing dynamics of the tether during the deployment, the details of which are drawn in Figue~\ref{fig:5}. For ST-SMC, the in-plane dynamic motion is regulated to vertical status at 4 $rad$, and the in-plane angle converges to the origin in the subsequent time. Hence the coalition of the length and the swing dynamics imply that the whole deployment mission is completed at about 4 $rad$ by utilizing ST-SMC. S-SMC makes the tether swing to the desired position at a low convergent speed near the vertical direction during 2 $rad$ to 4 $rad$.\par
The inputs in S-SMC for dynamic model~(\ref{eq:dynamic model}) are exhibited in Figure~\ref{fig:6}, in which it is obvious that the limited situations occur frequently. Thanks to the scalable dimensionless transformation determined conservatively, the tension exerted on the tether falls in the reasonable region.\par
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig3.eps}
\caption{Length dynamics of the deployment}
\label{fig:3}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig4.eps}
\caption{Length rate dynamics of the deployment}
\label{fig:4}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig5.eps}
\caption{Swing dynamics of the deployment}
\label{fig:5}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig6.eps}
\caption{Tension}
\label{fig:6}
\end{figure}
So far, the proposed control laws have been verified in the undisturbed environment. However, space perturbations and parametric uncertainties usually occur during the deployment of the TSS in engineering application. For checking the robustness of the proposed control laws, based on the dynamics~(\ref{eq:dynamics 1}), we have the following assumptions about uncertainties and disturbances:
\begin{align}\begin{split}
\frac{1}{\varpi}&\le\frac{\hat{\bar m}}{\bar m}\le \varpi\\
\frac{1}{\varpi}&\le\frac{\hat L}{L}\le \varpi\\
\frac{1}{\varpi}&\le\frac{\hat \Omega}{\Omega}\le \varpi\\
\bm\tau_d&=\left(0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000)\right)^T,
\end{split}\end{align}
where $\hat\bullet$ stands for the measurement of the term $\bullet$ with uncertainties. $\varpi$ is the uncertainty boundary parameter, and here we select $\varpi=2$. $\bm\tau_d$ is the external disturbance meaning  space perturbations, such as gravity gradient torque, aerodynamic interference and solar radiation pressure torque.  In this simulation, the measured terms can be specified as: $\hat{\bar m}=1.1\bar m$ may mean some space debris affixed to the subsatellite; $\bar L=1.02L$ and $\bar \Omega=0.98\Omega$ mean some measurement errors happening to the total length and the orbital angular velocity. In Fig~\ref{fig:6}, although the dynamics have already been  stable, the tension is switching in the high frequency. It is commonplace in the sliding mode control law design, called chattering, since the signum function is used in the input, which may cause unmodeled high frequency dynamics, and for reducing the chattering, we introduce the boundary layer technique into the proposed control laws with the thickness of the boundary layer specified as a positive scalar $\nu=1\times 10^{-4}$ in this simulation~\cite{1039802}.

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\end{figure}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}